% \documentclass[review=false, screen=true]{acmart}
\documentclass[format=acmsmall, review=false, screen=true]{acmart}

% Package to generate and customize Algorithm as per ACM style
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}

\usepackage{booktabs}
\usepackage{tabularx}

\usepackage{amsmath,amssymb,amsfonts}

\usepackage{enumitem}

\SetKwRepeat{Do}{do}{while}

\usepackage{lmodern}
\usepackage{courier}

\usepackage{listings}
\lstset{
  language=Java,
  backgroundcolor=\color{white},
  basicstyle=\ttfamily\small,
  breakatwhitespace=false,
  breaklines=true,
  captionpos=b,
  commentstyle=\color{red},
  frame=single,
  keepspaces=true,
  keywordstyle=\color{blue},
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  stringstyle=\color{brown},
  tabsize=2,
}

\usepackage{setspace}



% Document starts
\begin{document}
% Title portion. Note the short title for running heads
\title[COSC 3P91 -- Assignment 3]{COSC 3P91 -- Assignment 3 -- Village War Strategy Game}

\author{Group Member 1}
\affiliation{%
  \institution{Brock University}
  \streetaddress{1812 Sir Isaac Brock Way}
  \city{St. Catharines}
  \state{ON}
  \postcode{L2S 3A1}
  \country{Canada}
}

\begin{abstract}
This document describes the redesign and implementation of a console-based village
war strategy game for COSC 3P91 Assignment 3.  Building on Assignment 2, the
codebase has been restructured to apply three mandated Design Patterns: the
\textbf{Model--View--Controller (MVC)} architectural pattern, the
\textbf{Factory} creational pattern, and the \textbf{Adapter} structural pattern.
As a bonus, \textbf{XML persistence} (save/load) with a companion XSD schema is
also provided.  The game supports building, training, upgrading, NPC village
generation, attack exploration, and combat driven by the provided external
\texttt{ChallengeDecision} engine.  All Assignment 2 OOP concepts (generics,
local/anonymous classes, lambda expressions, custom exceptions, and Stream API
usage) are retained and still demonstrated in the redesigned codebase.
\end{abstract}

\maketitle

% ------------------------------------------------------------------
\section{Java Application Design and Implementation}
% ------------------------------------------------------------------

The game is implemented as a multi-package Java application.  The
Assignment~3 redesign introduces three additional packages and restructures
\texttt{Main} into a thin MVC wiring point:

\begin{itemize}[noitemsep]
  \item \texttt{game} -- model layer: \texttt{GameEngine}, \texttt{Village},
        \texttt{Army}, \texttt{Player}, \texttt{Defence}, \texttt{CollectResources},
        \texttt{WallClock}, and the generic \texttt{Repository<T>}.
  \item \texttt{gameelements} -- model layer: buildings, habitants, and resources
        (abstract hierarchies \texttt{Building}, \texttt{Habitant}/\texttt{Fighter}/\texttt{Peasant},
        \texttt{Resource}, and all concrete sub-classes).
  \item \textbf{\texttt{controller}} -- \texttt{GameController}: the MVC Controller.
  \item \textbf{\texttt{factory}} -- \texttt{BuildingFactory}, \texttt{HabitantFactory}: Factory pattern.
  \item \textbf{\texttt{persistence}} -- \texttt{VillageSerializer}: XML save/load bonus.
  \item \texttt{utility} -- \texttt{ChallengeDecisionAdapter} (Adapter pattern), \texttt{AttackOutcome},
        \texttt{GameMap}, \texttt{Position}, \texttt{Region}, legacy \texttt{Arbitrer}.
  \item \texttt{exceptions} -- four custom checked exceptions.
  \item \texttt{gui} -- \texttt{GraphicalInterface}: the MVC View.
  \item \texttt{ChallengeDecision} -- provided external combat-resolution engine (\emph{unmodified}).
\end{itemize}

\texttt{Main.java} is now a thin MVC composition root: it creates the Model,
View, and Controller, and enters the game loop.  All menu dispatch logic lives
in \texttt{GameController}.

% ------------------------------------------------------------------
\section{Design Patterns}
% ------------------------------------------------------------------

% ------------------------------------------------------------------
\subsection{MVC Architectural Pattern}
% ------------------------------------------------------------------

The codebase is divided into three cleanly separated layers (Table~\ref{tbl:mvc}).

\begin{table}[h]
  \centering
  \caption{MVC class mapping.}
  \label{tbl:mvc}
  \begin{tabularx}{\linewidth}{lX}
    \toprule
    \textbf{Layer} & \textbf{Classes / Responsibility} \\
    \midrule
    \textbf{Model} & \texttt{game/} and \texttt{gameelements/} -- all game state
                     and rules; no console I/O. \\
    \textbf{View}  & \texttt{gui.GraphicalInterface} -- all console rendering;
                     no game logic. \\
    \textbf{Controller} & \texttt{controller.GameController} -- reads user input
                     (forwarded from \texttt{Main}), interprets commands,
                     mutates the Model via its API, and delegates display
                     to the View. \\
    \textbf{Wiring} & \texttt{Main} -- creates M/V/C, enters event loop. \\
    \bottomrule
  \end{tabularx}
\end{table}

\texttt{GraphicalInterface} (View) exposes methods such as
\texttt{renderVillage()}, \texttt{renderBuildMenu()}, \texttt{renderTrainMenu()},
\texttt{renderUpgradeMenu()}, \texttt{renderTargets()}, and \texttt{renderArmy()}.
It receives data as parameters and never touches game logic.

\texttt{GameController.handleMenuChoice(String)} is the sole input-dispatch
method.  It calls Model API methods (e.g.\ \texttt{Village.build()},
\texttt{Army.attack()}) and forwards display work to the View.

\begin{spacing}{0.8}
\begin{lstlisting}[language=Java, caption={MVC wiring in Main.java.}, label=lst:mvc]
// Model
GameEngine engine = new GameEngine();
Player     player = new Player(1, villageName);

// View
GraphicalInterface view = new GraphicalInterface();

// Controller
GameController controller = new GameController(engine, player, view, scanner);

// Event loop
while (engine.isRunning()) {
    engine.tick();
    view.renderMenu();
    String input = scanner.nextLine().trim();
    controller.handleMenuChoice(input);
}
\end{lstlisting}
\end{spacing}

% ------------------------------------------------------------------
\subsection{Factory Creational Pattern}
% ------------------------------------------------------------------

Two factory classes centralise object instantiation so that the Controller
never uses \texttt{new} on concrete building or habitant classes directly
(Table~\ref{tbl:factory}).

\begin{table}[h]
  \centering
  \caption{Factory pattern roles.}
  \label{tbl:factory}
  \begin{tabularx}{\linewidth}{llX}
    \toprule
    \textbf{Factory} & \textbf{Client} & \textbf{Products} \\
    \midrule
    \texttt{BuildingFactory} & \texttt{GameController .buildBuilding()} &
      \texttt{Farm}, \texttt{GoldMine}, \texttt{IronMine}, \texttt{LumberMill},
      \texttt{ArcherTower}, \texttt{Cannon}, \texttt{VillageHall} \\
    \texttt{HabitantFactory} & \texttt{GameController .trainUnit()} &
      \texttt{Soldier}, \texttt{Archer}, \texttt{Knight}, \texttt{Catapult},
      \texttt{GoldMiner}, \texttt{IronMiner}, \texttt{Lumberman} \\
    \bottomrule
  \end{tabularx}
\end{table}

Both factories expose \texttt{create(String type)} and cost-query helpers
(\texttt{getGoldCost}, \texttt{getIronCost}, \texttt{getLumberCost}).
Adding a new building or unit type requires only a change inside the factory
class itself; the Controller is unaffected.

\begin{spacing}{0.8}
\begin{lstlisting}[language=Java, caption={Factory usage in GameController.}, label=lst:factory]
// BuildingFactory creates the correct Building sub-class
Building building = buildingFactory.create(type);
village.getBuildingsMutable().add(building);

// HabitantFactory creates the correct Habitant sub-class
Habitant habitant = habitantFactory.create(type);
player.getVillage().train(habitant, cg, ci, cl);
\end{lstlisting}
\end{spacing}

% ------------------------------------------------------------------
\subsection{Adapter Structural Pattern}
% ------------------------------------------------------------------

The provided \texttt{ChallengeDecision} package defines its own API
(\texttt{ChallengeEntitySet}, \texttt{ChallengeAttack}, \texttt{ChallengeDefense},
\texttt{ChallengeResource}, \texttt{ChallengeResult}) and \emph{must not be modified}.
The game model uses \texttt{Army} and \texttt{Village}.  The Object Adapter
\texttt{utility.ChallengeDecisionAdapter} bridges the two incompatible interfaces:

\begin{center}
\texttt{Army/Village} $\longrightarrow$ \texttt{ChallengeDecisionAdapter}
$\longrightarrow$ \texttt{ChallengeDecision.Arbitrer} $\longrightarrow$
\texttt{AttackOutcome}
\end{center}

\begin{itemize}[noitemsep]
  \item \textbf{Adaptee:} \texttt{ChallengeDecision.Arbitrer.challengeDecide()} -- external, unmodified.
  \item \textbf{Adapter:} \texttt{ChallengeDecisionAdapter.adapt(Army, Village)} -- maps each
        \texttt{Fighter} to a \texttt{ChallengeAttack}, each \texttt{Building} to a
        \texttt{ChallengeDefense}, resources to \texttt{ChallengeResource}; calls the external
        arbitrer; converts the \texttt{ChallengeResult} back to \texttt{AttackOutcome}.
  \item \textbf{Client:} \texttt{game.Army.attack()} -- calls the adapter instead of the
        old hand-rolled \texttt{utility.Arbitrer}.
\end{itemize}

\begin{spacing}{0.8}
\begin{lstlisting}[language=Java, caption={Adapter usage in Army.attack().}, label=lst:adapter]
// game.Army.attack() -- client uses the adapter
ChallengeDecisionAdapter adapter = new ChallengeDecisionAdapter();
AttackOutcome outcome = adapter.adapt(this, defender);

// Inside ChallengeDecisionAdapter.adapt() -- translation
for (Fighter f : attacker.getFighters()) {
    ChallengeAttack<Double,Double> ca =
        new ChallengeAttack<>(f.damage(), f.getHitPoints());
    attackList.add(ca);
}
// ... build defenseList and resourceList ...
ChallengeResult result =
    ChallengeDecision.Arbitrer.challengeDecide(challenger, challengee);
// ... convert ChallengeResult -> AttackOutcome ...
\end{lstlisting}
\end{spacing}

% ------------------------------------------------------------------
\subsection{XML Persistence (Bonus +15)}
% ------------------------------------------------------------------

\texttt{persistence.VillageSerializer} provides two static methods:

\begin{itemize}[noitemsep]
  \item \texttt{save(Village, String filePath)} -- serialises village name,
        resources, buildings (type, level, hit-points), and habitants (type)
        to a well-formed XML file using \texttt{javax.xml} APIs.
  \item \texttt{load(String filePath)} -- parses the XML file and reconstructs
        a \texttt{Village}, advancing building levels as needed.  XXE injection
        is prevented by disabling DOCTYPE declarations on the
        \texttt{DocumentBuilderFactory}.
\end{itemize}

The companion file \texttt{village\_schema.xsd} defines the XML Schema.  It
constrains building and habitant types via enumerations, restricts resource
values to \texttt{xs:double}, and limits building levels to positive integers.
The menu keys \texttt{s} (save) and \texttt{l} (load) in \texttt{GameController}
trigger these operations.

% ------------------------------------------------------------------
\section{Game Features}
% ------------------------------------------------------------------

\subsection{Building}

A player selects a building type from the build menu in
\texttt{GameController.buildBuilding()}, which uses \texttt{BuildingFactory}
to create the building and \texttt{Village.build()} to add it (enforcing the
20-building limit and deducting resources).

% ------------------------------------------------------------------
\subsection{Training}
% ------------------------------------------------------------------

Military and worker units are trained through \texttt{GameController.trainUnit()}.
\texttt{HabitantFactory.create(type)} instantiates the unit; \texttt{Village.train()}
deducts resources and appends the habitant.  Fighters can also be added to the army
via \texttt{Army.addFighter()} (capped at 10 units).

% ------------------------------------------------------------------
\subsection{Upgrading}
% ------------------------------------------------------------------

\texttt{Village.upgrade(Building)} upgrades any building already present in the
village.  It enforces two constraints:
\begin{itemize}[noitemsep]
  \item The building's level must be below the \texttt{VillageHall} level (the
        hall therefore acts as a global upgrade gate).
  \item Resources must cover the upgrade cost returned by
        \texttt{getUpgradeCostGold/Iron/Lumber()}.
\end{itemize}
After passing all checks, \texttt{Building.upgrade()} increments the level,
adds 50 hit-points, and calls the abstract hook \texttt{applyUpgradeBonus()}.

% ------------------------------------------------------------------
\subsection{Generating Villages}
% ------------------------------------------------------------------

\texttt{GameEngine.generateNpcVillages(int count)} creates \textit{count} NPC
villages stored in \texttt{Repository<Village>}.  A local class \texttt{NpcBuilder}
encapsulates construction logic; difficulty (1--3) is chosen randomly.

% ------------------------------------------------------------------
\subsection{Attack Exploring}
% ------------------------------------------------------------------

\texttt{GameEngine.getAvailableTargets()} returns NPC villages sorted by defence
score ascending (anonymous \texttt{Comparator}).  The player browses the list in
\texttt{GameController.exploreTargets()} and selects a target.

% ------------------------------------------------------------------
\subsection{Attacking}
% ------------------------------------------------------------------

\texttt{GameController.attackVillage()} calls \texttt{Army.attack(Village defender)},
which invokes the \texttt{ChallengeDecisionAdapter} (Adapter pattern) to delegate
to the provided external arbitrer.  Attack score is
\(\text{totalDamage} \times \sqrt{|\text{fighters}|}\).
On success, loot is added to the player's village; on failure, $\approx30$\,\%
of fighters are lost.

% ------------------------------------------------------------------
\subsection{Game Engine}
% ------------------------------------------------------------------

\texttt{GameEngine} owns the \texttt{WallClock} (tick-based time), the player and
NPC village \texttt{Repository} instances, and the \texttt{GameMap}.
The leaderboard sorts players by score with a method reference:
\texttt{Comparator.comparingInt(Player::getScore).reversed()}.

\begin{table}[h]
  \centering
  \caption{Building types, starting hit-points, and upgrade costs.}
  \begin{tabular}{lrrrr}
    \toprule
    \textbf{Building} & \textbf{HP} & \textbf{Gold} & \textbf{Iron} & \textbf{Lumber} \\
    \midrule
    Village Hall   & 500 & 100 & 50 & 50 \\
    Farm           & 200 & 25  & 15 & 10 \\
    Gold Mine      & 300 & 50  & 25 & 25 \\
    Iron Mine      & 300 & 40  & 30 & 20 \\
    Lumber Mill    & 300 & 15  & 5  & 30 \\
    Archer Tower   & 350 & 30  & 20 & 15 \\
    Cannon         & 400 & 50  & 40 & 20 \\
    \bottomrule
  \end{tabular}
\end{table}

% ------------------------------------------------------------------
\section{Object-Oriented Principles and Java Concepts}
% ------------------------------------------------------------------

% ------------------------------------------------------------------
\subsection{Generics}
% ------------------------------------------------------------------

The class \texttt{Repository<T>} (Listing~\ref{lst:repository}) is the
canonical demonstration of generics.  It is instantiated twice in
\texttt{GameEngine}: once as \texttt{Repository<Player>} and once as
\texttt{Repository<Village>}, giving strongly-typed access to each store
without raw-type casts.  The backing data structure is a \texttt{TreeMap<String, T>},
preserving lexicographic key order.

The method \texttt{Army.addAll(List<? extends Fighter>)} demonstrates an
upper-bounded wildcard: any list whose element type is \texttt{Fighter} or a
subtype is accepted, enabling reuse with heterogeneous fighter lists without
sacrificing type safety.

\begin{spacing}{0.8}
\begin{lstlisting}[language=Java, caption={Generic Repository (excerpt).}, label=lst:repository]
public class Repository<T> {
    private final TreeMap<String, T> store = new TreeMap<>();

    public void put(String key, T entity) { store.put(key, entity); }

    public List<T> findWhere(Predicate<T> predicate) {
        return store.values().stream()
                .filter(predicate)
                .collect(Collectors.toList());
    }
    public void forEach(Consumer<T> action) {
        store.values().forEach(action);
    }
}
\end{lstlisting}
\end{spacing}

% ------------------------------------------------------------------
\subsection{Local and Anonymous Classes}
% ------------------------------------------------------------------

\textbf{Local class.}~\texttt{GameEngine.generateNpcVillages()} declares a local
class \texttt{NpcBuilder} whose sole \texttt{build()} method constructs and
populates an NPC village.  Scoping the class locally keeps the builder logic
cohesive and prevents it from polluting the package namespace.

A second local class \texttt{BuildingMerger} appears inside
\texttt{Defence.getAllDefenceBuildings()}: it merges the archer-tower and
cannon lists into a single \texttt{List<Building>} using an upper-bounded
wildcard parameter.

\textbf{Anonymous class.}~\texttt{GameEngine.getAvailableTargets()} passes an
anonymous \texttt{Comparator<Village>} to \texttt{List.sort()} for ordering
targets by ascending defence score.  Similarly,
\texttt{Village.getBuildingsSortedByLevel()} uses an anonymous
\texttt{Comparator<Building>} to sort by descending level.

\begin{spacing}{0.8}
\begin{lstlisting}[language=Java, caption={Local class NpcBuilder and anonymous Comparator.}, label=lst:localanon]
// Local class inside generateNpcVillages()
class NpcBuilder {
    Village build(String name, int difficulty) {
        Village v = new Village(name, 200 + difficulty * 100, ...);
        if (difficulty >= 1) v.getBuildingsMutable().add(new ArcherTower());
        ...
        return v;
    }
}

// Anonymous Comparator inside getAvailableTargets()
targets.sort(new Comparator<Village>() {
    @Override
    public int compare(Village a, Village b) {
        return Double.compare(a.getDefenceScore(), b.getDefenceScore());
    }
});
\end{lstlisting}
\end{spacing}

% ------------------------------------------------------------------
\subsection{Lambda Expressions and Method References}
% ------------------------------------------------------------------

Lambdas and method references appear throughout the codebase:

\begin{itemize}[noitemsep]
  \item \texttt{Village.getDefenceScore()} -- \texttt{mapToDouble(b -> b.getLevel() * b.getHitPoints()).sum()}.
  \item \texttt{Village.getAttackScore()} -- stream pipeline with a filter lambda and a \texttt{mapToDouble} lambda.
  \item \texttt{Village.getStatusSummary()} -- \texttt{buildings.forEach(b -> sb.append(...))} and \texttt{habitants.forEach(h -> ...)}.
  \item \texttt{Army.recalculateDamage()} -- \texttt{mapToDouble(Fighter::damage).sum()}, using a method reference.
  \item \texttt{Army.getSummary()} -- \texttt{Collectors.groupingBy(f -> f.getClass().getSimpleName(), Collectors.counting())} followed by \texttt{forEach((type, count) -> ...)}.
  \item \texttt{GameEngine.getLeaderboard()} -- \texttt{sorted(Comparator.comparingInt(Player::getScore).reversed())}, a method reference composed with a reversal combinator.
  \item \texttt{CollectResources.collect()} -- \texttt{map(Peasant::work)} and subsequent \texttt{filter}/\texttt{mapToDouble} pipelines for each building type.
  \item \texttt{Defence.syncWithVillage()} -- \texttt{forEach(archerTowers::add)} and \texttt{forEach(cannons::add)}, instance method references.
\end{itemize}

\begin{spacing}{0.8}
\begin{lstlisting}[language=Java, caption={Lambda and method reference examples.}, label=lst:lambda]
// Village.getDefenceScore()
return buildings.stream()
        .mapToDouble(b -> b.getLevel() * b.getHitPoints())
        .sum();

// Army.recalculateDamage() -- method reference
damage = fighters.stream()
        .mapToDouble(Fighter::damage)
        .sum();

// GameEngine.getLeaderboard() -- method reference + combinator
return playerRepository.getAll().stream()
        .sorted(Comparator.comparingInt(Player::getScore).reversed())
        .collect(Collectors.toList());
\end{lstlisting}
\end{spacing}

% ------------------------------------------------------------------
\subsection{Custom Exceptions}
% ------------------------------------------------------------------

Four checked exceptions in the \texttt{exceptions} package model domain-specific
error conditions:

\begin{itemize}[noitemsep]
  \item \texttt{InsufficientResourcesException} -- thrown by
        \texttt{Village.checkResources()} when gold, iron, or lumber is short.
        Stores the resource type, required amount, and available amount.
  \item \texttt{BuildingLimitExceededException} -- thrown by \texttt{Village.build()}
        when the 20-building cap is reached.  Stores current count and cap.
  \item \texttt{MaxLevelReachedException} -- thrown by \texttt{Building.upgrade()}
        when the level-5 cap is reached, or by \texttt{Village.upgrade()} when
        the VillageHall gate is hit.
  \item \texttt{InvalidOperationException} -- a general-purpose exception for
        illegal state transitions (e.g.\ attacking with an empty army,
        removing a non-existent building).
\end{itemize}

Each exception class provides two constructors: a single-string constructor for
programmatic use and a structured constructor that generates a descriptive message
from the relevant parameters.  The \texttt{Main} game loop catches all four
exceptions and displays a user-friendly error line prefixed with \texttt{[!]}.

\begin{spacing}{0.8}
\begin{lstlisting}[language=Java, caption={InsufficientResourcesException.}, label=lst:exception]
public class InsufficientResourcesException extends Exception {
    private final String resourceType;
    private final double required;
    private final double available;

    public InsufficientResourcesException(
            String resourceType, double required, double available) {
        super(String.format(
            "Insufficient %s: required %.1f but only %.1f available.",
            resourceType, required, available));
        ...
    }
}
\end{lstlisting}
\end{spacing}

% ------------------------------------------------------------------
\subsection{Java Utility Classes (Collections)}
% ------------------------------------------------------------------

The following standard collection types are used:

\begin{itemize}[noitemsep]
  \item \texttt{ArrayList<Building>}, \texttt{ArrayList<Habitant>},
        \texttt{ArrayList<Fighter>} -- the primary mutable collections inside
        \texttt{Village} and \texttt{Army}.
  \item \texttt{TreeMap<String, T>} -- the backing store of \texttt{Repository<T>},
        chosen for its guaranteed sorted key order and $O(\log n)$ access.
  \item \texttt{Collections.unmodifiableList()} -- used by \texttt{Village.getBuildings()}
        and \texttt{Village.getHabitants()} to expose read-only views of internal
        state, protecting encapsulation.
  \item \texttt{Map<String, Long>} produced by
        \texttt{Collectors.groupingBy(..., Collectors.counting())} inside
        \texttt{Army.getSummary()} to count fighters by type.
\end{itemize}

% ------------------------------------------------------------------
\subsection{I/O and Streams}
% ------------------------------------------------------------------

The Stream API drives much of the game's data processing.  Key uses:

\begin{itemize}[noitemsep]
  \item \textbf{Filtering:}  \texttt{Village.getFighters()} and
        \texttt{Village.getPeasants()} filter the \texttt{habitants} list by
        \texttt{instanceof} predicates.
  \item \textbf{Mapping:}  \texttt{CollectResources.collect()} calls
        \texttt{map(Peasant::work)} to transform each peasant into the
        \texttt{Resource} it produces.
  \item \textbf{Aggregation (reduce / sum):}
        \texttt{Defence.getTotalDamage()} accumulates archer-tower and cannon
        damage via \texttt{mapToDouble(...).reduce(0, Double::sum)}.
  \item \textbf{Grouping:}  \texttt{Army.getSummary()} groups fighters by
        class name and counts them with \texttt{Collectors.groupingBy}.
  \item \textbf{Sorting:}  \texttt{Village.getFightersSortedByDamage()} returns
        fighters ordered by descending damage using a composed comparator:
        \texttt{Comparator.comparingDouble(Fighter::damage).reversed()}.
\end{itemize}

Console I/O uses a \texttt{Scanner} wrapping a \texttt{BufferedReader} around
\texttt{System.in} in \texttt{Main.main()}, giving buffered line-oriented input
with minimal latency.

\begin{spacing}{0.8}
\begin{lstlisting}[language=Java, caption={Stream-based resource collection in CollectResources.}, label=lst:streams]
List<Resource> produced = peasants.stream()
        .map(Peasant::work)
        .collect(Collectors.toList());

totalGold += buildings.stream()
        .filter(b -> b instanceof GoldMine)
        .mapToDouble(b -> ((GoldMine) b).getGoldProd())
        .sum();
\end{lstlisting}
\end{spacing}

% ------------------------------------------------------------------
\subsection{Performance Analysis}
% ------------------------------------------------------------------

The game runs in a tight console loop.  Each iteration is $O(B + H)$ where $B$
is the number of buildings and $H$ the number of habitants, both bounded by
constants ($B \le 20$, $H \le 10$ fighters plus unlimited peasants in
practice).  The \texttt{Repository<T>} uses a \texttt{TreeMap} giving
$O(\log n)$ key-based lookup; for the small NPC village set ($n = 5$ by default)
this is negligible.  Combat resolution in \texttt{Arbitrer} is $O(1)$.

% ------------------------------------------------------------------
\subsection{Conclusions}
% ------------------------------------------------------------------

The Assignment~3 redesign applies three Design Patterns on top of the complete
Assignment~2 game:
\begin{itemize}[noitemsep]
  \item \textbf{MVC} -- clean separation of Model (\texttt{game/}, \texttt{gameelements/}),
        View (\texttt{gui.GraphicalInterface}), and Controller (\texttt{controller.GameController}).
  \item \textbf{Factory} -- \texttt{BuildingFactory} and \texttt{HabitantFactory} centralise
        all object instantiation for buildings and habitants.
  \item \textbf{Adapter} -- \texttt{ChallengeDecisionAdapter} bridges the game model to the
        provided external \texttt{ChallengeDecision} engine without modification.
  \item \textbf{XML Persistence (bonus)} -- \texttt{VillageSerializer} saves/loads village
        state to/from \texttt{village\_save.xml}, validated by \texttt{village\_schema.xsd}.
\end{itemize}
All six game features and all Assignment~2 OOP concepts are retained in the redesigned codebase.


\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{acmsmall-sample-bibfile}


\appendix

% ------------------------------------------------------------------
\section{Main.java -- MVC Wiring and Game Loop (excerpt)}
% ------------------------------------------------------------------
\begin{spacing}{0.8}
\begin{lstlisting}[language=Java]
public static void main(String[] args) {
    Scanner scanner = new Scanner(new BufferedReader(
                        new InputStreamReader(System.in)));
    // View (MVC)
    GraphicalInterface view = new GraphicalInterface();
    // Model (MVC)
    GameEngine engine = new GameEngine();
    Player     player = new Player(1, villageName);
    engine.registerPlayer(player);
    engine.generateNpcVillages(5);
    engine.setRunning(true);
    // Controller (MVC)
    GameController controller = new GameController(engine, player, view, scanner);

    while (engine.isRunning()) {
        engine.tick();
        view.renderMenu();          // View renders
        String input = scanner.nextLine().trim();
        try {
            controller.handleMenuChoice(input); // Controller dispatches
        } catch (InsufficientResourcesException e) {
            System.out.println("  [!] " + e.getMessage());
        } catch (MaxLevelReachedException e) {
            System.out.println("  [!] " + e.getMessage());
        } catch (BuildingLimitExceededException e) {
            System.out.println("  [!] " + e.getMessage());
        } catch (InvalidOperationException e) {
            System.out.println("  [!] " + e.getMessage());
        }
    }
}
\end{lstlisting}
\end{spacing}

\clearpage

% ------------------------------------------------------------------
\section{Village.java -- Score Calculations (excerpt)}
% ------------------------------------------------------------------
\begin{spacing}{0.8}
\begin{lstlisting}[language=Java]
// Defence score: sum of level * hitPoints over all buildings
public double getDefenceScore() {
    return buildings.stream()
            .mapToDouble(b -> b.getLevel() * b.getHitPoints())
            .sum();
}

// Attack score: sum of damage * level for each Fighter habitant
public double getAttackScore() {
    return habitants.stream()
            .filter(h -> h instanceof Fighter)
            .map(h -> (Fighter) h)
            .mapToDouble(f -> f.damage() * f.getLevel())
            .sum();
}

// Buildings sorted by level (anonymous Comparator)
public List<Building> getBuildingsSortedByLevel() {
    List<Building> sorted = new ArrayList<>(buildings);
    Collections.sort(sorted, new Comparator<Building>() {
        @Override
        public int compare(Building a, Building b) {
            return Integer.compare(b.getLevel(), a.getLevel());
        }
    });
    return sorted;
}
\end{lstlisting}
\end{spacing}

\clearpage

% ------------------------------------------------------------------
\section{Army.java -- Attack and Wildcard (excerpt)}
% ------------------------------------------------------------------
\begin{spacing}{0.8}
\begin{lstlisting}[language=Java]
// Attack score with squad-size synergy
public double getAttackScore() {
    if (fighters.isEmpty()) return 0;
    return damage * Math.sqrt(fighters.size());
}

// Bounded wildcard: accepts any List<? extends Fighter>
public void addAll(List<? extends Fighter> newFighters)
        throws InvalidOperationException {
    for (Fighter f : newFighters) addFighter(f);
}

// Method reference: Fighter::damage
private void recalculateDamage() {
    damage = fighters.stream()
            .mapToDouble(Fighter::damage)
            .sum();
}
\end{lstlisting}
\end{spacing}

\end{document}
